### 2015-11-18 pldi-2016/paper.tex
changes

\documentclass[10pt,preprint,blind,clearpagebib]{sigplanconf}

\usepackage{listings}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage[svgnames]{xcolor}
\definecolor{lcl}{RGB}{140,0,100}
\usepackage[colorlinks=true,breaklinks,draft=false]{hyperref}
\hypersetup{urlcolor=lcl,linkcolor=lcl,citecolor=lcl}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{graphics} 
\usepackage{stmaryrd}
\usepackage{amsthm}
\usepackage{ftnright}
\usepackage[T1]{fontenc}
\usepackage{semantic}
\usepackage{enumitem}
\usepackage[hang,flushmargin]{footmisc} 
\usepackage[notquote]{hanging}
\usepackage{flushend}

% =================================================================================================

% TODO: Discuss 'stability' of the inference (minor change in schema does not cause dramatic cahnges)
% TODO: Say that variant is a last-resort thing
% TODO: Practical evaluation (lots of downloads, lots of Github - how often people customize samples?)
% TODO: Explain why we need option type (because T + null is a variant)
% TODO: In the PLDI version, we delete all the unnecessary runtime conversions
% but we should say that those happen for practical reasons (e.g. float -> int, or anyhing to string)

\urlstyle{sf}

\makeatletter
\renewcommand{\@makefntext}[1]{%
  \parindent 1em%
  \raggedright
  \begin{hangparas}{0.8em}{1}
  \noindent {$^{\@thefnmark}$~#1}
  \end{hangparas}
}
\makeatother

\newcommand{\langl}{\begin{picture}(4.5,7)
\put(1.1,2.5){\rotatebox{60}{\line(1,0){5.5}}}
\put(1.1,2.5){\rotatebox{300}{\line(1,0){5.5}}}
\end{picture}}
\newcommand{\rangl}{\begin{picture}(4.5,7)
\put(.9,2.5){\rotatebox{120}{\line(1,0){5.5}}}
\put(.9,2.5){\rotatebox{240}{\line(1,0){5.5}}}
\end{picture}}

\newcommand{\lang}{\begin{picture}(5,7)
\put(1.1,2.5){\rotatebox{45}{\line(1,0){6.0}}}
\put(1.1,2.5){\rotatebox{315}{\line(1,0){6.0}}}
\end{picture}}
\newcommand{\rang}{\begin{picture}(5,7)
\put(.1,2.5){\rotatebox{135}{\line(1,0){6.0}}}
\put(.1,2.5){\rotatebox{225}{\line(1,0){6.0}}}
\end{picture}}

\newcommand{\llangl}{\langl\hspace{-0.35em}\langl}
\newcommand{\rrangl}{\rangl\hspace{-0.35em}\rangl}

\definecolor{cmtclr}{rgb}{0.0,0.6,0.0}
\definecolor{numclr}{rgb}{0.0,0.4,0.0}
\definecolor{kvdclr}{rgb}{0.0,0.0,0.6}
\definecolor{strclr}{rgb}{0.5,0.1,0.0}
\definecolor{prepclr}{rgb}{0.0,0.0,0.0}

\newcommand{\kvd}[1]{\textnormal{\textcolor{kvdclr}{\sffamily #1}}}
\newcommand{\num}[1]{\textnormal{\textcolor{numclr}{\sffamily #1}}}
\newcommand{\str}[1]{\textnormal{\textcolor{strclr}{\sffamily "#1"}}}
\newcommand{\strf}[1]{\textnormal{\textcolor{strclr}{\sffamily #1}}}
\newcommand{\ident}[1]{\textnormal{\sffamily #1}}
\newcommand{\lident}[1]{\textnormal{\sffamily\`{}\hspace{-0.25em}\`{}\hspace{-0.1em}#1\`{}\hspace{-0.25em}\`{}}}
\newcommand{\cmt}[1]{\textit{\sffamily\textcolor{cmtclr}{#1}}}

\newcommand{\lsep}[0]{\;\; | \;\;}
\newcommand{\narrow}[1]{\hspace{-0.7em} #1 \hspace{-0.7em}}

\newcommand{\tsep}[0]{\; \triangledown \;}
\newcommand{\tytag}{\ident{tag}}
\newcommand{\dropopt}[1]{\lfloor#1\rfloor}
\newcommand{\addopt}[1]{\lceil#1\rceil}
\newcommand{\tytagof}{\ident{tagof}}
\newcommand{\nameoftag}{\ident{nameof}}

\newcommand{\reduce}{\rightsquigarrow}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\semalt}[1]{\llangl #1 \rrangl}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

% =================================================================================================

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}
\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
%\doi{nnnnnnn.nnnnnnn}

%\titlebanner{Unpublished draft, March 2015}        % These are ignored unless
%\preprintfooter{short description of paper}        % 'preprint' option specified.

\title{F\# Data: \textnormal{Making structured data first-class citizens}}
%\subtitle{Subtitle Text, if any}

\authorinfo{Tomas Petricek}
           {University of Cambridge}
           {tomas@tomasp.net}
\authorinfo{Gustavo Guerra}
           {Microsoft Corporation, London}
           {gustavo@codebeside.org}
\authorinfo{Don Syme}
           {Microsoft Research, Cambridge}
           {dsyme@microsoft.com}
\maketitle

% =================================================================================================

\begin{abstract}
Most modern applications interact with external services and access data in structured formats such 
as XML, JSON and CSV. Static type systems do not understand such formats, often making data access
more cumbersome. Should we give up and leave the messy world of external data to dynamic typing 
and runtime checks? Of course, not!

In this paper, we integrate external structured data into the F\# type system. As most real-world data
do not come with an explicit schema, we develop a type inference algorithm that infers a type from 
representative sample documents and integrate it into the F\# type system using type providers.

Our library significantly reduces the amount of code that developers need to write when 
accessing data. It also provides additional safety guarantees, arguably, as much as possible if 
we abandon the closed world assumption.
\end{abstract}

\category{D.3.3}{Programming Languages}{Language Constructs and Features }
\keywords F\#, Type Providers, Inference, JSON, XML

% =================================================================================================
%
%   ###                                                                       
%    #  #    # ##### #####   ####  #####  #    #  ####  ##### #  ####  #    # 
%    #  ##   #   #   #    # #    # #    # #    # #    #   #   # #    # ##   # 
%    #  # #  #   #   #    # #    # #    # #    # #        #   # #    # # #  # 
%    #  #  # #   #   #####  #    # #    # #    # #        #   # #    # #  # # 
%    #  #   ##   #   #   #  #    # #    # #    # #    #   #   # #    # #   ## 
%   ### #    #   #   #    #  ####  #####   ####   ####    #   #  ####  #    # 
%
% =================================================================================================

\section{Introduction}
\label{sec:introduction}

Social network clients, applications for finding tomorrow's weather or searching train schedules
all communicate with one or more services over the network and display the resulting information. 
Increasing number of such services provide REST-based end-points that return data as CSV, XML
or JSON. Despite numerous schematization efforts, most services do not come with an 
explicit schema. At best, the documentation provides sample responses for typical requests.

For example, OpenWeatherMap provides an end-point for getting current weather for a 
given city\footnote{See ``Current weather data'': \url{http://openweathermap.org/current}}. 
The documentation shows one sample to illustrate the typical response. Using standard 
JSON and web libraries, we might write:
%
\begin{equation*}
\begin{array}{l}
 \kvd{let}~\ident{doc}=\ident{Http.Request}(\str{http://api.owm.org/?q=NYC}) \\
 \kvd{match}~\ident{JsonValue.Parse}(\ident{doc})~\kvd{with} \\
 |~\ident{Record}(\ident{root})\rightarrow \\
 \quad \kvd{match}~\ident{Map.find}~\str{main}~\ident{root}~\kvd{with} \\
 \quad |~\ident{Record}(\ident{main})\rightarrow \\
 \quad \quad \kvd{match}~\ident{Map.find}~\str{temp}~\ident{main}~\kvd{with} \\
 \quad \quad |~\ident{Number}(\ident{num})\rightarrow \ident{printfn}~\str{Lovely \%f!}~\ident{num} \\
 \quad \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\
 \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\
 |~\_\rightarrow \ident{failwith}~\str{Incorrect format} 
\end{array}
\end{equation*}
%
The code assumes that the response has a particular format described in the documentation. The
root node must be a record with a \str{main} field, which has to be another record containing
a numerical \str{temp} field. When the format is incorrect, the data access simply fails
with an exception.

Using the JSON type provider from F\# Data, we can write code with exactly the 
same functionality in two lines:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{W} = \ident{JsonProvider}\langl\str{http://api.owm.org/?q=NYC}\rangl \\[0.1em]
 \ident{printfn}~\str{Lovely \%f!}~(\ident{W.GetSample().Main.Temp})
\end{array}
\end{equation*}
%
$\ident{JsonProvider}\langl\str{...}\rangl$ invokes a type provider at 
compile-time with the URL as a sample. The type provider infers the structure of the response
and provides a type with a \ident{GetSample} method that returns a parsed JSON with nested
properties \ident{Main.Temp}, returning the temperature as a number.

The rest of the paper describes the mechanism and discusses its safety properties. 
The key novel contributions are:

\begin{itemize}
\item We present F\# Data type providers for XML, CSV and JSON (\S\ref{sec:providers}) 
  and practical aspects of their implementation that contributed to their industrial 
  adoption (\S\ref{sec:impl}).

\item We describe a predictable type inference algorithm for structured data formats, 
  based on a \emph{preferred typing} relation, that underlies the type providers 
  (\S\ref{sec:inference}).

\item We give a formal model (\S\ref{sec:formal}) and use it to prove
  \emph{relativized type safety} for the type providers (\S\ref{sec:safety}).
  This adapts the ML-style type safety for the context of the web.

\item The supplementary screencast show developer experience using 
  F\# Data with JSON, XML and CSV.
\end{itemize}

% =================================================================================================
%
%   #######                                                                                  
%      #    #   # #####  ######    #####  #####   ####  #    # # #####  ###### #####   ####  
%      #     # #  #    # #         #    # #    # #    # #    # # #    # #      #    # #      
%      #      #   #    # #####     #    # #    # #    # #    # # #    # #####  #    #  ####  
%      #      #   #####  #         #####  #####  #    # #    # # #    # #      #####       # 
%      #      #   #      #         #      #   #  #    #  #  #  # #    # #      #   #  #    # 
%      #      #   #      ######    #      #    #  ####    ##   # #####  ###### #    #  ####  
%
% =================================================================================================

\section{Structural type providers}
\label{sec:providers}

We start with an informal overview that shows how F\# Data type providers simplify working with 
JSON and XML. We introduce the necessary aspects of F\# type providers along the 
way. The examples in this section illustrate a number of key properties of our type inference 
algorithm:

\begin{itemize}
\item Our mechanism is predictable. The user directly works with the inferred types and should 
  understand why a specific type was inferred from a given sample.\footnote{In particular, we do 
  not use probabilistic methods where adding an additional sample could completely change the 
  shape of the type.}

\item Our inference mechanism prefers records. This supports developer tooling --
  most F\# editors provide code completion hints on ``.'' and so types with properties (records)
  are easier to use than types that require pattern matching.

\item Finally, we handle a number of practical concerns important in the real world. This includes 
  support for different numerical types, \kvd{null} values and missing data.
\end{itemize}

% --------------------------------------------------------------------------------------------------

\subsection{Working with JSON documents}
\label{sec:providers-json}

The JSON format used in \S\ref{sec:introduction} is a popular data exchange format based on 
JavaScript data structures. The following is the definition of \ident{JsonValue} 
used earlier to represent JSON:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{JsonValue} = \\[0.1em]
 \quad|~ \ident{Number}~\kvd{of}~\ident{decimal} ~~~~|~ \ident{Boolean}~\kvd{of}~\ident{bool} \\[0.1em]
 \quad|~ \ident{String}~\kvd{of}~\ident{string} ~~\qquad\, |~ \ident{Null} \\[0.1em]
 \quad|~ \ident{Record}~\kvd{of}~\ident{Map}\langl\ident{string}, \ident{JsonValue}\rangl \\[0.1em]
 \quad|~ \ident{Array}~\kvd{of}~\ident{JsonValue}[] \\[0.1em]
\end{array}
\end{equation*}
%
The earlier example used only a nested record containing a number. To demonstrate other 
aspects of the JSON type provider, we look at an example that also involves an array:
%
{\small{
\begin{verbatim}
  [ { "name":"Jan", "age":25 }, { "name":"Tomas" },
    { "name":"Alexander", "age":3.5 } ]
\end{verbatim}
}}
%
\noindent
Say we want to print the names and an age if it is available. The standard approach would 
be to pattern match on the parsed \ident{JsonValue}, check that the top-level node is a \ident{Array}
and iterate over the elements checking that each element is a \ident{Record} with certain properties. 
We would throw an exception for values in an incorrect format. Even with helper functions, 
we would still need to specify field names as strings, which is error prone and can not be 
statically checked.

Assuming \strf{people.json} is the above example and \ident{data} is a string containing
JSON in the same format, we can write:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{People}~=~\ident{JsonProvider}\langl\str{people.json}\rangl\hspace{1em} \\[0.6em]
 \kvd{for}~\ident{item}~\kvd{in}~\ident{People.Parse}(\ident{data})~\kvd{do}\\[0.1em]
 \quad\ident{printf}~\str{\%s }~\ident{item.Name}\\[0.1em]
 \quad\ident{Option.iter}~(\ident{printf}~\str{(\%f)})~\ident{item.Age}
\end{array}
\end{equation*}
%
In contrast to the earlier example, we now use a local file \strf{people.json} as a sample for 
the type inference, but then processes data from another source. The code achieves the same 
simplicity as when using dynamically typed languages, but it is statically type-checked.

\paragraph{Type providers.}
The notation $\ident{JsonProvider}\langl\str{people.json}\rangl$ passes a \emph{static parameter} 
to the type provider. Static parameters are resolved at compile-time and have to be constant. 
The provider analyzes the sample and generates a type  \ident{People}. Most F\# editors also 
execute the type provider in the background at development-time and use the provided types 
in auto-completion.

The \ident{JsonProvider} uses a type inference algorithm (\S\ref{sec:inference})  and
infers the following types from the sample:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Entity}~=  \\
 \quad \kvd{member}~\ident{Name}~:~\ident{string} \\
 \quad \kvd{member}~\ident{Age}~:~\ident{option}\langl \ident{decimal}\rangl \\[0.5em]
 \kvd{type}~\ident{People}~=  \\
 \quad \kvd{member}~\ident{GetSample}~:~\ident{unit}~\rightarrow~\ident{Entity}[] \\
 \quad \kvd{member}~\ident{Parse}~:~\ident{string}~\rightarrow~\ident{Entity}[] \\
\end{array}
\end{equation*}
%
The type \ident{Entity} represents the person. The field \ident{Name} is available for all
sample values and is inferred as \ident{string}. The field \ident{Age} is marked as optional,
because the value is missing in one sample. The two age values are an integer $25$ and a 
decimal $3.5$ and so the common inferred type is \ident{decimal}.

The type \ident{People} has two methods for reading data. \ident{GetSample} parses the
sample used for the inference and \ident{Parse} parses a JSON string. This lets us read
data at runtime, provided that it is in the same format as the static sample.

\paragraph{Error handling.}
In addition to the structure of the types, the type provider also specifies what code should be 
executed at run-time in place of \ident{item.Name} and other operations. The runtime behaviour is 
the same as in the earlier hand-written sample (\S\ref{sec:introduction}) -- a member access 
throws an exception if the document does not have the expected format.

Informally, the safety property (\S\ref{sec:safety}) states that if the inputs
are subtypes of some of the provided samples (i.e.~the samples are representative), then
no exceptions will occur. In other words, we cannot avoid all failures, but we can prevent some.
Moreover, if the OpenWeatherMap changes the response format, the sample in \S\ref{sec:introduction}
will not re-compile and the user knows that the code needs to be changed.

\paragraph{The role of records.}
The sample code is easy to write thanks to the fact that most F\# editors provide code completion
when ``.'' is typed (see the supplementary screencast). The developer does not need to look at the 
sample JSON file to see what fields are available. To support this scenario, our inference algorithm 
prefers records (we also treat XML elements and CSV rows as records) and only infers types that 
require pattern matching as the last resort.

In the above example, this is demonstrated by the fact that \ident{Age} is marked as optional.
An alternative is to provide two different record types (one with \ident{Name} and other with 
\ident{Name} and \ident{Age}), but this would complicate the processing code.

% -------------------------------------------------------------------------------------------------

\subsection{Processing XML documents}
\label{sec:providers-xml}

XML documents are formed by nested elements with attributes. We can view elements as records with 
a field for each attribute and an additional special field for the nested contents (which is a 
collection of elements).

Consider a simple extensible document format where a root element
{\ttfamily\small <doc>} can contain a number of document element, one of which is
{\ttfamily\small <heading>} representing headings:
%
{\small{
\begin{verbatim}
  <doc>
    <heading>Working with JSON</heading>
    <p>Type providers make this easy.</p>
    <heading>Working with XML</heading>
    <p>Processing XML is as easy as JSON.</p>
    <image source="xml.png" />
  </doc>
\end{verbatim}
}}
%
\noindent
The F\# Data library has been designed primarilly to simplify reading of data. For example,
say we want to print all headings in the document. The sample shows a part of the document structure 
(in particular the {\ttfamily\small <heading>} element), but it does not show all possible elements 
(say, {\ttfamily\small <table>}). Assuming the above document is \strf{sample.xml}, we can write:
%
\noindent
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Document}~=~\ident{XmlProvider}\langl\str{sample.xml}\rangl\hspace{1em} \\[0.5em]
 \kvd{let}~\ident{root}~=~\ident{Document.Load}(\str{c:/pldi/another.xml})\\
 \kvd{for}~\ident{elem}~\kvd{in}~\ident{root.Doc}~\kvd{do}\\
 \quad\ident{Option.iter}~(\ident{printf}~\str{ - \%s})~\ident{elem.Heading}\\
\end{array}
\end{equation*}
%
The example iterates over a collection of elements returned by \ident{root.Doc}. The type of element 
provides a typed access to elements known from the sample and so we can write \ident{elem.Heading}, 
which returns an optional string value.

\paragraph{Open world and variant types.}
We do not infer a union type representing a choice between heading, paragraph and image. By its 
nature, XML is extensible and the sample cannot include all possible nodes.\footnote{Even when the 
document structure is defined using XML Schema, documents may contain elements prefixed with other 
namespaces.} The type of \ident{elem} is thus a top type (we call it a \emph{variant}) annotated 
with the statically known possibilities. It is mapped to the following F\# type:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Element}~=  \\[0.1em]
 \quad \kvd{member}~\ident{Heading}~:~\ident{option}\langl \ident{string} \rangl\\
 \quad \kvd{member}~\ident{Paragraph}~:~\ident{option}\langl \ident{string} \rangl\\
 \quad \kvd{member}~\ident{Image}~:~\ident{option}\langl \ident{Image} \rangl\\
\end{array}
\end{equation*}
%
The type provides a nice access to the statically known elements, but it encodes the 
\emph{open-world assumption}. The user also needs to handle the case when the value is none 
of the statically known elements.

As a consequence, the variant type can be seen as the top type. The above code will work
correctly on document that contains any elements, including those not listed in the sample. This
is extremely useful when reading data from external sources, because the sample document only
needs to contain values relevant to the application. It is also worth noting that our type 
inference attempts to minimze the number of variant types that appear in the resulting type --
a variant is not used if there is another option (see Corollary~\ref{thm:no-unions} and 
\S\ref{sec:impl-hetero}).

% -------------------------------------------------------------------------------------------------

\subsection{Summary}
\label{sec:providers-sum}
In addition to type providers for JSON and XM, F\# Data also implements a type provider for CSV 
(\S\ref{sec:impl-parsing}). We treat CSV files as lists of records (with field for each column) 
and so CSV is handled directly by our inference algorithm.

Throughout the introduction, we used data sets that demonstrate typical issues that are frequent
in the real-world (missing data, inconsistent encoding of primitive values and heterogeneous 
structures). In our experience, these are the most common issues. The following JSON response with 
government debt information returned by the World Bank API\footnote{Available at 
\url{http://data.worldbank.org}} demonstrates all three problems:
%
{\small{
\begin{verbatim}
  [ { "page": 1, "pages": 5 },
    [ { "indicator": "GC.DOD.TOTL.GD.ZS",
        "date": "2012", "value": null },
      { "indicator": "GC.DOD.TOTL.GD.ZS",
        "date": "2010", "value": "35.14229" } ] ]
\end{verbatim}
}}
%
\noindent
First of all, the top-level element is a collection containing two values of different kind.
The first is a record with meta-data about the current page and the second is an array with data. 
The actual F\# Data implementation supports a concept of heterogeneous collections (briefly outlined
in \S\ref{sec:impl-hetero}) and provides a type with properties \ident{Record} for the former and 
\ident{Array} for the latter. Second, the \ident{value} field is \kvd{null} for some records. Third, 
numbers can be represented in JSON as numeric literals (without quotes), but here, they are 
returned as string literals instead.\footnote{This is often used to avoid non-standard numerical 
types of JavaScript.}

% =================================================================================================
%
%   #######                                                                                  
%      #    #   # #####  ######    # #    # ###### ###### #####  ###### #    #  ####  ###### 
%      #     # #  #    # #         # ##   # #      #      #    # #      ##   # #    # #      
%      #      #   #    # #####     # # #  # #####  #####  #    # #####  # #  # #      #####  
%      #      #   #####  #         # #  # # #      #      #####  #      #  # # #      #      
%      #      #   #      #         # #   ## #      #      #   #  #      #   ## #    # #      
%      #      #   #      ######    # #    # #      ###### #    # ###### #    #  ####  ###### 
%
% =================================================================================================

\section{Inferring Types from Data}
\label{sec:inference}

Our type inference algorithm for structured formats is based on a subtyping relation. When 
inferring the type, we infer the most specific types of individual values (CSV rows, JSON or XML 
nodes) and recursively find a preferred typing of all child nodes or all sample documents.

We first define the type of structured data $\sigma$. Note that this type is distinct 
from the programming language types $\tau$ (type providers generate the latter from the former). 
Next, we define the subtyping relation on structural types $\sigma$ and describe the algorithm 
for finding a preferred typing.

\subsection{Inferred Types}
\label{sec:inference-types}

We distinguish between \emph{non-nullable types} that always have a valid value (written as 
$\hat{\sigma}$) and \emph{nullable types} that encompass missing and \kvd{null} values 
(written as $\sigma$). We write $\nu$ for record field names and $\nu_?$ for names that can be empty:
%
\begin{equation*}
\begin{array}{rcl}
 \hat{\sigma} &\narrow{=}& \nu_?\; \{ \nu_1 : \sigma_1, \ldots, \nu_n : \sigma_n \} \\[0.1em]
                &\narrow{|}& \ident{float} \lsep \ident{decimal} \lsep \ident{int} \lsep \ident{bool} \lsep \ident{string} 
 \\[0.6em] 
       \sigma &\narrow{=}& ~\hat{\sigma}~ \lsep \kvd{option}\langl \hat{\sigma} \rangl \lsep [\sigma] \\[0.1em]
              &\narrow{|}& \kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl \lsep ~\bot~ \lsep \kvd{null}
\end{array}
\end{equation*}

\noindent
Non-nullable types include records (consisting of an optional name and fields with their types) 
and primitives. Records arising from XML documents are named, while records used by 
JSON and CSV providers are unnamed.

The three numerical primitives, \ident{int} for integers, \ident{decimal} for small precise decimal 
numbers and \ident{float} for floating-point numbers, are related in a preference order as discussed in 
\S\ref{sec:inference-subtyping}.

Any non-nullable type can be wrapped in the \kvd{nullable} constructor to explicitly permit the 
\kvd{null} value. Later, type providers will map \kvd{nullable} to the standard F\# option type. A collection $[\sigma]$ 
is also nullable: \kvd{null} are treated as empty collections, and missing fields of collection type
evaluate to empty collections. The type 
$\kvd{null}$ is inhabited by the $\kvd{null}$ value (using an overloaded notation) and $\bot$ 
is the bottom type.

A variant type \kvd{any} is annotated with the types it may represent in the sample document.
As discussed earlier (\S\ref{sec:providers-xml}), variants force the user to handle the case when 
a value is of a different type than the statically known ones and so it can be seen as a family of 
top types. For the same reason, variants also implicitly permit the \kvd{null} value.

% --------------------------------------------------------------------------------------------------

\begin{figure}
\begin{center}
\includegraphics[scale=0.80,trim=5mm 5mm 5mm 5mm,clip]{images/hierarchy.pdf} % left bottom right top
\end{center}
\vspace{-0.5em}
\caption{Preference relation between types inferred from data samples}
\label{fig:subtyping-diagram}
\vspace{-0.5em}
\end{figure}

% -------------------------------------------------------------------------------------------------

\subsection{Preference relation}
\label{sec:inference-subtyping}

Figure~\ref{fig:subtyping-diagram} provides a basic intuition about the relationship between
possible types inferred from data samples. The upper part shows non-nullable types (with records and primitive types) and 
the lower part shows nullable types with \kvd{null}, collections and optional values. We 
abbreviate $\kvd{option}\langl\sigma\rangl$ as $\sigma?$ and the diagram omits links between the two parts;
any type $\hat{\sigma}$ is preferred to $\kvd{option}\langl\hat{\sigma}\rangl$.

\begin{definition}
We write $\sigma_1 \leq \sigma_2$ to denote that type $\sigma_2$ is preferred over $\sigma_1$. The 
relation is defined as a transitive reflexive closure of the following rules:

\noindent
\begin{align}
  \label{eq:sub-prim}\tag{P1}
  \ident{float}\,\leq\,\ident{decimal}\,&\leq\,\ident{int}&\\[-0.2em]
  \label{eq:sub-null}\tag{P2}
  \sigma &\leq \kvd{null}  &(\textnormal{iff}~\sigma \neq \hat{\sigma})  \\[-0.2em]
  \label{eq:sub-opt}\tag{P3}
  \kvd{option}\langl\hat{\sigma}\rangl &\leq \hat{\sigma}  &(\textnormal{for all}~\hat{\sigma})\\[-0.2em]
  \label{eq:sub-opt-cov}\tag{P4}
  \kvd{option}\langl\hat{\sigma_1}\rangl &\leq 
    \kvd{option}\langl\hat{\sigma_2}\rangl  &(\textnormal{if}~\hat{\sigma_1} \leq \hat{\sigma_2})\\[-0.2em]
  \label{eq:sub-col}\tag{P5}
  [\sigma_1] &\leq [\sigma_2]  &(\textnormal{if}~\sigma_1 \leq \sigma_2) \\[-0.2em]
  \label{eq:sub-bot}\tag{P6}
  \sigma &\leq \bot  &(\textnormal{for all}~\sigma)\\[-0.2em]
  \label{eq:sub-var-top}\tag{P7}
  \kvd{any}\langl \sigma_1, \ldots, \sigma_n\rangl &\leq \sigma 
\end{align}

\noindent
\begin{align}
\label{eq:sub-record1}\tag{R1}
\begin{array}{l}
 \nu_?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \} \leq \\
 \nu_?~\{ \nu_1\!:\!\sigma_1', .., \nu_n\!:\!\sigma_n' \}
\end{array} \qquad ~~~~(\textnormal{if}~\sigma_i \leq \sigma_i')
\end{align}
\vspace{-1em}
\begin{align}
\label{eq:sub-record2}\tag{R2}
\begin{array}{l}
 \nu_?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \} \leq \\
 \nu_?~\{ \nu_1\!:\!\sigma_1, .., .., \nu_m\!:\!\sigma_m \}
\end{array} \qquad (\textnormal{when}~m \geq n)
\end{align}
\vspace{-1em}
\begin{align}
\label{eq:sub-record3}\tag{R3}
\begin{array}{l}
 \nu_?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \} \leq \\
 \nu_?~\{ \nu_{\pi(1)}\!:\!\sigma_{\pi(1)}, .., \nu_{\pi(m)}\!:\!\sigma_{\pi(m)} \}
\end{array} ~~(\pi~\textnormal{perm.})
\end{align}
\vspace{-1em}
\begin{align}
\label{eq:sub-record4}\tag{R4}
\begin{array}{l}
 \nu_?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n, \nu_{n+1}\!:\!\kvd{null} \} \leq \\
 \nu_?~\{ \nu_1\!:\!\sigma_1, .., \nu_n\!:\!\sigma_n \}
\end{array}
\end{align}
\end{definition}

% --------------------------------------------------------------------------------------------------

\begin{figure*}[t]
\noindent
\begin{equation*}
\boxed{
\begin{array}{c}
\begin{array}{rclcl}
 \tytag &\narrow{=}& \ident{bool} \\
        &\narrow{\lsep}& \ident{string} &\narrow{\lsep}& \ident{number}  \\
        &\narrow{\lsep}& \ident{any}  &\narrow{\lsep}& \ident{collection} \\
        &\narrow{\lsep}& \ident{record} &\narrow{\lsep}& \ident{named}\;~\nu 
\end{array}
%
\quad\;\;\;
%
\begin{array}{rcl}
 \tytagof(\ident{string}) &\narrow{=}& \ident{string}\\
 \tytagof(\ident{bool}) &\narrow{=}& \ident{bool}\\
 \tytagof([\sigma]) &\narrow{=}& \ident{collection}\\
 \tytagof(\hat{\sigma}~\kvd{option}) &\narrow{=}& \tytagof(\hat{\sigma})
\end{array}
%
\;\;
%
\begin{array}{rcl}
 \tytagof(\kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl) &\narrow{=}& \ident{any}\\
 \tytagof(\{ \nu_1 : \sigma_1, \; \ldots \; , \nu_n : \sigma_n \}) &\narrow{=}& \ident{rec-anon}\\
 \tytagof(\nu\; \{ \nu_1 : \sigma_1, \; \ldots \; , \nu_n : \sigma_n \}) &\narrow{=}& \ident{rec-named}~\nu \\
 \multicolumn{3}{c}{ \tytagof(\sigma) = \ident{number}\quad \sigma \in\{ \ident{int},\ident{decimal},\ident{float} \}  } 
\end{array}
%
\\[3em]
%
\begin{array}{rcll}
 \addopt{\hat{\sigma}} &\narrow{=}& \hat{\sigma}~\kvd{option} &(\textnormal{non-nullable types})\\
 \addopt{\sigma} &\narrow{=}& \sigma &(\textnormal{otherwise})
\end{array}
\qquad\qquad
\begin{array}{rcll}
 \dropopt{\hat{\sigma}~\kvd{option}} &\narrow{=}& \hat{\sigma} &(\textnormal{option})\\
 \dropopt{\sigma} &\narrow{=}& \sigma &(\textnormal{otherwise})
\end{array}
\end{array}
}
\end{equation*}

\begin{equation*}
\hspace{3em}
\inference[(record-1)\;]
  { (\nu_i = \nu'_j) \Leftrightarrow (i = j) \wedge (i \leq k)
      \qquad \forall i\in\{ 1 .. k \}.(\sigma_i \tsep \sigma'_i \vdash \sigma''_i) }
  { \begin{array}{l}
    \nu_? \; \{ \nu_1 \!:\! \sigma_1,  \; \ldots \;, \nu_k \!:\! \sigma_k, \; \ldots \;, \nu_n \!:\! \tau_n \} \tsep
    \nu_? \; \{ \nu'_1 \!:\! \sigma'_1, \; \ldots \;, \nu'_k \!:\! \sigma'_k, \; \ldots \;, \nu'_m \!:\! \tau'_m \} \vdash\\
    \nu_? \; \{ \nu_1 \!:\! \sigma''_1, \; \ldots \; , \nu_k \!:\! \sigma''_k, 
                            \nu_{k+1} \!:\! \addopt{\sigma_{k+1}}, \ldots, \nu_n \!:\! \addopt{\sigma_n},
                            \nu'_{k+1} \!:\! \addopt{\sigma'_{k+1}}, \ldots, \nu'_m \!:\! \addopt{\sigma'_m} \}
    \end{array} }
\end{equation*}
\vspace{-2em}

% union with something else
\begin{equation*}
\textnormal{\footnotesize{(var-1)}}\;\;
\inference
  {\exists i . \tytagof(\sigma_i) = \tytagof(\sigma) & \sigma \tsep \sigma_i \vdash \sigma_i' & \tytagof(\sigma)\neq\ident{any}}
  {\sigma \tsep \kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl \vdash 
    \kvd{any}\langl\sigma_1, \ldots, \dropopt{\sigma_i'}, \ldots, \sigma_n\rangl}
\;\;
\inference
  {\nexists i . \tytagof(\sigma_i) = \tytagof(\sigma) & \tytagof(\sigma)\neq\ident{any}}
  {\sigma \tsep \kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl \vdash 
    \kvd{any}\langl\sigma_1, \ldots, \sigma_n, \dropopt{\sigma}\rangl}
\end{equation*}
\vspace{-2em}

% two union types
\begin{equation*}
\begin{array}{l}
\inference[(var-2)\;]
  { (\forall i\in \{ 1 .. k \})\quad \tytagof(\sigma_i) = \tytagof(\sigma'_i) \quad 
    \sigma_i \tsep \sigma_i' \vdash \sigma''_i}
  { \begin{array}{l}
    \kvd{any}\langl \sigma_1, \ldots, \sigma_k,  \ldots, \sigma_n\rangl \tsep 
    \kvd{any}\langl \sigma'_1, \ldots, \sigma'_k, \ldots, \sigma'_m\rangl \vdash\\
    \kvd{any}\langl \sigma''_1, \ldots, \sigma''_k, \sigma_{k+1}, \ldots, \sigma_{n}, \sigma'_{k+1}, \ldots, \sigma'_{m}\rangl
    \end{array} }
\end{array}    
\begin{array}{l}
\inference[(var-3)\;]
  {(\forall i\in\{1,2\}) & \tytagof(\sigma_1) \neq \tytagof(\sigma_2) \\ 
   \tytagof(\sigma_i) \neq \ident{any} & \nexists\sigma_i' . (\sigma_i = \kvd{option}\langl\sigma_i'\rangl)  }
  {\sigma_1 \tsep \sigma_2 \vdash \kvd{any}\langle\dropopt{\sigma_1}, \dropopt{\sigma_2}\rangle}
\end{array}  
\end{equation*}
\vspace{-2em}

\begin{equation*}
\inference[(order-1)\;]
  { \nu_? \; \{ \nu_1 : \sigma_1, \ldots, \nu_n : \sigma_n \} \tsep \sigma \vdash \sigma' }
  { \nu_? \; \{ \nu_{\pi(1)} : \sigma_{\pi(1)}, \ldots, \nu_{\pi(n)} : \sigma_{\pi(n)} \} \tsep \sigma \vdash \sigma' }
\quad
\inference[(order-2)\;]
  { \kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl \tsep \sigma \vdash \sigma' }
  { \kvd{any}\langl\sigma_{\pi(1)}, \ldots, \sigma_{\pi(n)}\rangl \tsep \sigma \vdash \sigma' }
\quad (\pi~\textnormal{perm.})  
\end{equation*}
\vspace{-2em}

\begin{equation*}
\hspace{3em}
% option types
\inference[(opt)\;]
  {\hat{\sigma_1} \tsep \sigma_2 \vdash \sigma}
  {\kvd{option}\langl\hat{\sigma_1}\rangl \tsep \sigma_2 \vdash \kvd{option}\langl\dropopt{\sigma}\rangl}
\qquad
% primitive types
\inference[(prim)\;]
  {\sigma_1 \leq \sigma_2 &
   \tytagof(\sigma_1) = \tytagof(\sigma_2) = \ident{number} }
  {\sigma_1 \tsep \sigma_2 \vdash \sigma_1}\quad
\end{equation*}
\vspace{-2em}

% all rules are symmetric and reflexive
\begin{equation*}
\inference[(list)\;]
  {\sigma_1 \tsep \sigma_2 \vdash \sigma}
  {[\sigma_1] \tsep [\sigma_2] \vdash [\sigma]}
\quad
\inference[(sym)\;]
  {\sigma_1 \tsep \sigma_2 \vdash \sigma}
  {\sigma_2 \tsep \sigma_1 \vdash \sigma}
%
\quad
\begin{array}{l}
 \textnormal{\footnotesize{(refl)}}\;\; \sigma \tsep \sigma \vdash \sigma\\[0.6em]
 % bottom type
 \textnormal{\footnotesize{(bot)}}\;\; \bot \tsep \sigma \vdash \sigma
\end{array}
%
\quad
% null and nullable
 \begin{array}{l}
 \textnormal{\footnotesize{(null-1)}}\;\; \sigma \tsep \kvd{null} \vdash \sigma \quad(\sigma \leq \kvd{null}) \\[0.6em]
 \textnormal{\footnotesize{(null-2)}}\;\; \sigma \tsep \kvd{null} \vdash \sigma~\kvd{option} \quad(\sigma :\ngtr \kvd{null})
 \end{array}
\end{equation*}

\caption{Inference judgements that define the most preferred typing relation}
\label{fig:subtyping-cst}
\end{figure*}

% --------------------------------------------------------------------------------------------------

\noindent
Here is a summary of the key aspects of the definition:
\begin{itemize}
\item Numeric types (\ref{eq:sub-prim}) are ordered by the range they represent;
  \ident{int} is a 32-bit integer, \ident{decimal} has a range of $-1^{29}$ to $1^{29}$ and 
  \ident{float} is a floating-point number (but imprecise).

\item The \kvd{null} type preferred to all nullable types (\ref{eq:sub-null}), i.e. all 
  types excluding non-nullable types $\hat{\sigma}$. Any non-nullable type is a 
  preferred to its nullable version (\ref{eq:sub-opt}) and both options and collections are 
  covariant (\ref{eq:sub-opt-cov}, \ref{eq:sub-col}).

\item There is a bottom type (\ref{eq:sub-bot}) and variants behave as top types, because
  any type $\sigma$ is a subtype of any variant (\ref{eq:sub-var-top}). There is a range of
  variants (no \emph{single} top type), but any variant is a subtype of any other variant.

\item As usual, the subtyping on records is covariant (\ref{eq:sub-record1}), subtype can have 
  additional fields (\ref{eq:sub-record2}) and fields can  be reordered (\ref{eq:sub-record3}). 
  The interesting rule is (\ref{eq:sub-record4}). It implies that subtype can omit fields, 
  provided that \kvd{null} is a valid value for the field (i.e. the field is optional).
\end{itemize}

\noindent
If we were type-checking programs, defining variants as a family of top types (which are all 
mutual subtypes) would not add any value over having a single top type, but our purpose is 
different. We focus on reading structured data and, as shown earlier (\S\ref{sec:providers-xml}), 
the types forming a variant provide a valuable information about the sample document.

A particularly important rule is (\ref{eq:sub-record4}), which allows subtype to have fewer record 
elements. This allows us to prefer records over variants. For example, given 
$\{ \ident{name}\!:\!\ident{string} \}$ and $\{ \ident{name}\!:\!\ident{string}, \ident{age}\!:\!\ident{int} \}$,
we find a preferred typing $\{ \ident{name}\!:\!\ident{string}, \ident{age}\!:\!\ident{int}~\kvd{option} \}$.
This is a least upper bound (clarified below) and more usable than another preferred typing 
$\kvd{var}\langl \{ \ident{name}\!:\!\ident{string} \}, \{ \ident{name}\!:\!\ident{string}, \ident{age}\!:\!\ident{int} \}\rangl$.

Some of the aspects of our system (numeric types and handling of missing data) offer a range of
choices. For example, OCaml also has different numerical types and would likely always handle 
\kvd{null} explicitly.

% -------------------------------------------------------------------------------------------------

\begin{figure*}
\noindent
\begin{equation*}
\begin{array}{l}
\ident{asFloat}(i) \,\reduce f\quad(f = i)\\
\ident{asFloat}(d) \reduce f\quad(f = d)\\
\ident{asFloat}(f) \reduce f
\\[0.5em]
\ident{asDec}(i) \,\reduce d\quad(d = i)\\
\ident{asDec}(d) \reduce d
\end{array}
\qquad
\begin{array}{l}
\ident{hasType}(\nu_? \{ \nu_1 \!:\! \sigma_1, \ldots, \nu_n \!:\! \sigma_n \}, \nu'_? \{ \nu'_1\mapsto s_1, \ldots, \nu'_m\mapsto s_m \}) \reduce (\nu_? = \nu'_?) ~\wedge \\
  \quad (~ ((\nu_1 = \nu'_1) \wedge \ident{hasType}(\sigma_1, s_1)) \vee\ldots\vee ((\nu_1 = \nu'_m) \wedge \ident{hasType}(\sigma_1, s_m)) \vee \ldots \vee\\
  \quad ~\; ((\nu_n = \nu'_1) \wedge \ident{hasType}(\sigma_n, s_1)) \vee\ldots\vee ((\nu_n = \nu'_m) \wedge \ident{hasType}(\sigma_n, s_m))~)
\\[0.5em]
\ident{hasType}([\sigma], [s_1; \ldots; s_n]) \reduce \ident{hasType}(\sigma, s_1)\wedge\ldots\wedge\ident{hasType}(\sigma, s_n) \\
\ident{hasType}([\sigma], \kvd{null}) \reduce \kvd{true} \\  
\end{array}  
\end{equation*}
%
\vspace{-0.5em}
%
\begin{equation*}
\begin{array}{l}
\ident{isNull}(\kvd{null}) \reduce \kvd{true} \\
\ident{isNull}(\_) \reduce \kvd{false} 
\\[0.5em]
\ident{getField}(\nu,\nu_i, \nu~\{\ldots, \nu_i=s_i, \ldots\}) \reduce s_i\\
\ident{getField}(\bullet, \nu_i, \{\ldots, \nu_i=s_i, \ldots\}) \reduce s_i\\
\ident{getField}(\nu,\nu', \nu~\{\ldots, \nu_i=s_i, \ldots\}) \reduce \kvd{null}\quad(\nexists i.\nu_i=\nu' )\\
\ident{getField}(\bullet, \nu', \{\ldots, \nu_i=s_i, \ldots\}) \reduce \kvd{null}\quad(\nexists i.\nu_i=\nu' )
\\[0.5em]
\ident{getChildren}([s_1; \ldots; s_n]) \reduce [s_1; \ldots; s_n]  \\  
\ident{getChildren}(\kvd{null}) \reduce [~] 
\end{array}
~
\begin{array}{l}
\ident{hasType}(\ident{string}, t) \reduce \kvd{true} \\
\ident{hasType}(\ident{bool}, s) \reduce \kvd{true} \quad(\textnormal{when}~s\in{\kvd{true},\kvd{false}} )\\
\ident{hasType}(\ident{float}, s) \reduce \kvd{true} \quad(\textnormal{when}~s=i, s=d ~\textnormal{or}~ s=f) \\
\ident{hasType}(\ident{decimal}, s) \reduce \kvd{true} \quad(\textnormal{when}~s=i ~\textnormal{or}~ s=f) \\
\ident{hasType}(\ident{int}, i) \reduce \kvd{true}\\
\ident{hasType}(\kvd{option}\langl\sigma\rangl, s) \reduce \ident{hasType}(\sigma,s) \\
\ident{hasType}(\kvd{option}\langl\sigma\rangl, s) \reduce \kvd{true} \\
\ident{hasType}(\kvd{any}\langl\ldots\rangl, s) \reduce \kvd{true} \\
\ident{hasType}(\_, \_) \reduce \kvd{false} \\
\end{array}
\end{equation*}

\caption{Reduction rules for conversion functions}
\label{fig:op-conversions}
\vspace{-0.5em}
\end{figure*}

% -------------------------------------------------------------------------------------------------

\subsection{Most preferred typing relation}
\label{sec:inference-commonsuper}

Given two typings, the \emph{most preferred typing} relation finds a greatest lower bound
(Theorem~\ref{thm:lub}). When possible, it avoids variant types and prefers records 
(Corollary~\ref{thm:no-unions}), which is important for usability as discussed earlier
(\S\ref{sec:providers-json}).

\begin{definition}
A \emph{preferred typing} of $\sigma_1$ and $\sigma_2$ is a typing $\sigma$, written 
$\sigma_1 \triangledown \sigma_2 \vdash \sigma$, obtained according to the inference rules in 
Figure~\ref{fig:subtyping-cst}.
\end{definition}

\noindent
When finding a preferred typing of records (\emph{record-1}), we return a record that has the 
union of their fields. The types of shared fields become preferred typings of their respective 
types while fields that are present in only one record are marked as optional. The (\emph{order-1})
rule lets us reorder fields.

Although all variants are mutual subtypes, we find one that best represents the sample. We avoid
nested variants and limit the number of cases. This is done by grouping types 
that have a preferred typing which is not a variant. For example, rather than inferring 
$\kvd{any}\langl\ident{int}, \kvd{any}\langl\ident{bool}, \ident{decimal}\rangl\rangl$, 
our algorithm finds the preferred typing of $\ident{int}$ and $\ident{decimal}$ and produces 
$\kvd{any}\langl\ident{decimal},\ident{bool}\rangl$. To identify types that have a preferred typing 
which is not a variant, we group the types by a tag. The tag of a type is obtained using a function 
\tytagof{(-)}. The function is not defined on $\bot$ and $\kvd{null}$ as those are hanled by the 
(\emph{bot}) and (\emph{null-1}) or (\emph{null-2}).

The handling of variants is specified using three rules. When combining a non-variant and a
variant (\emph{var-1}), the variant may or may not already contain a case with the tag of the other
type. If it does, the two types are combined, otherwise a new case is added.
When combining two variants (\emph{var-2}), we group the cases that have shared tags.
Finally, (\emph{var-3}) covers the case when we are combining two distinct non-variant types.
As variants implicitly permit \kvd{null} values, we use an auxiliary function $\dropopt{-}$ to
make nullable types non-nullable (when possible) to simplify the type.

The remaining rules are straightforward. For collections and options, we find the preferred typing 
of the contained value(s); for compatible primitive types, we choose their supertype (\emph{prim})
and a preferred typing with \kvd{null} is either the type itself (\emph{null-1}) or an option 
(\emph{null-2}).

\paragraph{Properties.}
The partially ordered set of types does not have a \emph{unique} least upper bound, but it does 
have a least upper bound with respect to an equivalence between mutual subtypes. The common 
supertype relation is well-defined (Theorem~\ref{thm:func}) and finds the least upper bound 
(Theorem~\ref{thm:lub}).

\begin{definition}
Types $\sigma_1$, $\sigma_2$ are \emph{equivalent}, written $\sigma_1 \equiv \sigma_2$ iff
they are mutual subtypes, i.e. $\sigma_1 <: \sigma_2 \wedge \sigma_1 \leq \sigma_2$.
An equivalence class of a type $\sigma$ is $\mathcal{E}(\sigma) = \{ \sigma' \;|\; \sigma\equiv\sigma' \}$.
\end{definition}

\noindent
The equivalence class $\mathcal{E}$ sheds light on the structure of structural types. It defines
a lattice with bottom $\{ \bot \}$ and top (set of all variant types). It also joins records with
reordered fields (of same names and types) and types that contain other equivalent types (e.g.~a
lists containing variant values).

\begin{theorem}[Function]
\label{thm:func}
\raggedright
For all $\sigma_1$ and $\sigma_2$ there exists exactly one $\sigma$ such that 
$\sigma_1 \triangledown \sigma_2 \vdash \sigma$.
Furthermore, if $\sigma_1 \triangledown \sigma_2 \vdash \sigma'$
then it holds that $\sigma \leq \sigma'$ and $\sigma' \leq \sigma$.
\end{theorem}
\begin{proof}
The pre-conditions of rules in Figure~\ref{fig:subtyping-cst} are disjoint, with the exception 
of (\emph{order}), (\emph{sym}) and (\emph{refl}). These rules produce types that are subtypes
of each other and this property is preserved by rules that use $\triangledown$ recursively.
\end{proof}

\begin{theorem}[Least upper bound]
\label{thm:lub}
\raggedright
If $\sigma_1 \triangledown \sigma_2 \vdash \sigma$ then $\sigma$ is a least upper bound, i.e. 
$\sigma \leq \sigma_1$ and $\sigma \leq \sigma_2$ and for all $\sigma'$ such that $\sigma' \leq \sigma_1$
and $\sigma' \leq \sigma_2$, it holds that $\sigma' \leq \sigma$.
\end{theorem}
\begin{proof}
By induction over $\vdash$. Note that the algorithm never produces nested variant or nested
option. When one type is variant, the only preferred typing is also a variant (\emph{var-2}); 
for (\emph{var-3}) no other preferred typing exists because the two types have distinct 
tags and are not options.
\end{proof}

\begin{corollary}
\label{thm:no-unions}
Given $\sigma_1, \sigma_2, \sigma$ such that $\sigma \leq \sigma_1$ and $\sigma \leq \sigma_2$ and 
$\sigma$ is not a variant, then $\sigma_1 \tsep \sigma_2 \vdash \sigma'$ 
and $\sigma' \equiv \sigma$.
\end{corollary}
\begin{proof}
Consequence of Theorem~\ref{thm:lub} and the fact that the set of all variants is the
top element of $\mathcal{E}$.
\end{proof}

\subsection{Inferring types from values}
\label{sec:formal-inferval}

The preferred typing algorithm (\S\ref{sec:inference}) is at the core of the type inference 
for structural values, but we have not yet clarified how it is used. Given a JSON, XML or CSV 
document, F\# Data constructs a structural value $s$ from the sample. The following defines a mapping 
$\semalt{s_1,\ldots,s_n}$ which turns a collection of sample values $s_1, \ldots, s_n$ into a 
structural type $\sigma$:
%
\begin{equation*}
\begin{array}{rclcrclcrcl}
 \semalt{i} &\narrow{=}& \ident{int} && \semalt{d} &\narrow{=}& \ident{decimal} && \semalt{b} &\narrow{=}& \ident{bool}\\
 \semalt{f} &\narrow{=}& \ident{float} && \semalt{t} &\narrow{=}& \ident{string} && \semalt{\kvd{null}}  &\narrow{=}& \kvd{null}\\
\end{array}
\end{equation*}
\noindent
\vspace{-0.5em}
\begin{equation*}
\begin{array}{l}
 \semalt{s_1, \ldots, s_n} = \sigma_n \quad\textnormal{where}\\[0.1em]
 \qquad\sigma_0 = \bot,~ \forall i\in \{ 1.. n \}.~ \sigma_{i-1} \triangledown \semalt{s_i} \vdash \sigma_i
 \\[0.5em]
 \semalt{[s_1; \ldots; s_n]} = [\semalt{s_1, \ldots, s_n}]
 \\[0.5em]
 \semalt{\nu?~\{ \nu_1 \mapsto s_1, \ldots, \nu_n \mapsto s_n \}} =\\[0.1em]
 \qquad\nu?~\{ \nu_1:\semalt{s_1}, \ldots, \nu_n :\semalt{s_n} \}
\end{array}
\end{equation*}
%
We overload the notation and write $\semalt{s}$ when inferring type from a single value. Primitive 
values are mapped to their corresponding types. For records, we return a type with field types 
inferred based on individual values.

When infering a value from multiple samples, we use the preferred typing relation to find a common 
type for all values (starting with $\bot$). This operation is used at the top-level (when calling 
a type provider with multiple samples) and also when inferring the type of a collection.

% ==================================================================================================
% 
%    #######
%    #        ####  #####  #    #   ##   #      # ######   ##   ##### #  ####  #    #
%    #       #    # #    # ##  ##  #  #  #      #     #   #  #    #   # #    # ##   #
%    #####   #    # #    # # ## # #    # #      #    #   #    #   #   # #    # # #  #
%    #       #    # #####  #    # ###### #      #   #    ######   #   # #    # #  # #
%    #       #    # #   #  #    # #    # #      #  #     #    #   #   # #    # #   ##
%    #        ####  #    # #    # #    # ###### # ###### #    #   #   #  ####  #    #
%
% ==================================================================================================

\section{Formalising type providers}
\label{sec:formal}

In this section, we build the theoretical framework for proving relativised type safety of 
structural type providers. We discuss the runtime representation of documents and 
primitive operations (\S\ref{sec:formal-convert}), we embed these into a formal model of an F\# 
subset (\S\ref{sec:formal-ff}) and we describe how structural type providers turn
inferred structural types into F\# types (\S\ref{sec:formal-tp}).

% -------------------------------------------------------------------------------------------------

\subsection{Structural values and conversions}
\label{sec:formal-convert}

We represent JSON, XML and CSV documents using the same first-order \emph{structural values} defined
as follows:
%
\begin{equation*}
\begin{array}{lcl}
 s &\narrow{=}& i \lsep d \lsep f \lsep t \lsep \kvd{true} \lsep \kvd{false} \lsep \kvd{null} \\[0.1em]
   &\narrow{|}& [s_1; \ldots; s_n] \lsep \nu_?\{ \nu_1 \mapsto s_1, \ldots, \nu_n \mapsto s_n \}
\end{array}
\end{equation*}
%
The first few cases represent primitive values ($i$ for integers, $d$ for decimals, $f$ for floating
point numbers and $t$ for strings) and the \kvd{null} value. A collection is written as a 
list of values in square brackets. A record starts with an optional name $\nu_?$, followed by a 
sequence of field assignments $\nu_i \mapsto s_i$.

As indicated by the subtyping, our system permits certain runtime conversions (e.g.~an integer
\num{1} can be treated as a floating-point \num{1.0}). The following primitive operations 
implement the conversions and other helpers:
%
\begin{equation*}
\begin{array}{lcl}
 op  &\narrow{=}& \ident{asDec}(s) \lsep \ident{asFloat}(s) \lsep \ident{getChildren}(s) \\
     &\narrow{|}& \ident{getField}(\nu?, \nu, s) \lsep \ident{isNull}(s) \lsep \ident{hasType}(\sigma, s)
\end{array}
\end{equation*}
%
The operations are used in code generated by type providers. Their behavior is defined
by the rules in Figure~\ref{fig:op-conversions}. The conversion functions (\ident{asDec}
and \ident{asFloat}) only perform conversions required by the subtyping (our actual 
implementation is more lax and performs additional conversions).

The \ident{getField} operation is used to access a field of a record. The operation ensures that
the actual record name matches the expected name (we write $\bullet$ for the name of an unnamed 
record). The \ident{getChildren} operation returns elements of a list and turns \kvd{null} into 
an empty collction.

Finally, we also define two helper functions. The \ident{isNull} operation is used to check whether 
value is \kvd{null} and \ident{hasType} is a runtime type test. The most complex case is handling 
of records where we check that for each field $\nu_1, \ldots, \nu_n$ in the type, the actual record 
value has a field of the same name with a matching type. The last line defines a ``catch all'' 
pattern, which returns \kvd{false} for all remaining cases. Although not elegant, the rules 
can be expressed using just Featherweight F\#, discussed in the next section.

% -------------------------------------------------------------------------------------------------

\begin{figure}
\noindent
\begin{equation*}
\begin{array}{rl}
 \textnormal{\footnotesize{(member)}}&
 \hspace{-0.4em}
 \inference
 { \kvd{type}~C(\overline{x:\tau})=\ldots \kvd{member}~N_i : \tau_i = e_i \ldots }
 { (\kvd{new}~C(\overline{v})).N_i \reduce e_i[\overline{x} \leftarrow \overline{v}] }\\
 \\
 \textnormal{\footnotesize{(eq1)}}&
 v=v\reduce\kvd{true} \qquad \textnormal{\footnotesize{(eq2)}}~~v=v'\reduce\kvd{false}\\
 \\
 \textnormal{\footnotesize{(cond1)}}&
 \hspace{-0.4em}
 \kvd{if}~\kvd{true}~\kvd{then}~e_1~\kvd{else}~e_2 ~\reduce~ e_1 \\
 \\
 \textnormal{\footnotesize{(cond2)}}&
 \hspace{-0.4em}
 \kvd{if}~\kvd{false}~\kvd{then}~e_1~\kvd{else}~e_2 ~\reduce~ e_2 \\
 \\
 \textnormal{\footnotesize{(match1)}}&
 \hspace{-1em}
 \begin{array}{l}
  \kvd{match}~\ident{None}~\kvd{with} \\
  \ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_2 \\
 \\
 \textnormal{\footnotesize{(match2)}}&
 \hspace{-1em}
 \begin{array}{l}
    \kvd{match}~\ident{Some}(v)~\kvd{with} \\
    \ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_1[x\leftarrow v]\\
 \\
 \textnormal{\footnotesize{(match3)}}&
 \hspace{-1em}
 \begin{array}{l}
  \kvd{match}~[v_1;\ldots;v_n]~\kvd{with} \\[0em]
  [x_1;\ldots;x_m ] \rightarrow e_1 \,|\, \_ \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_2\quad(m\neq n) \\
 \\
 \textnormal{\footnotesize{(match4)}}&
 \hspace{-1em}
 \begin{array}{l}
  \kvd{match}~[v_1;\ldots;v_n]~\kvd{with} \\[0em]
  [x_1;\ldots;x_n ] \rightarrow e_1 \,|\, \_ \rightarrow e_2
 \end{array} \hspace{-0.5em} ~\reduce~ e_1[\overline{x}\leftarrow\overline{v}] \\
 \\
 \textnormal{\footnotesize{(map)}}&
 \hspace{-0.4em}
 \ident{map}~(\lambda x.e)~[v_1; \ldots] ~\reduce~ [e[x\leftarrow v_1]; \ldots] \\
 \\
 \textnormal{\footnotesize{(ctx)}}&
 \hspace{-0.4em}
  E[e] \reduce E[e'] \qquad\qquad(\textnormal{when}~e \reduce e')\\
\end{array}
\end{equation*}

\caption{Featherweight F\# -- Remaining reduction rules}
\label{fig:ff-reduction}
\end{figure}

% -------------------------------------------------------------------------------------------------

\subsection{Featherweight F\#}
\label{sec:formal-ff}

We now extend the semantics discussed so far and add a minimal subset of F\# needed to model F\# 
Data. Type providers use classes and so we focus on the F\# object model and combine aspects of SML 
\cite{sml} and Featherweight Java \cite{fwjava}.

We only need classes with parameter-less members and without inheritance. A class has a single 
implicit constructor and the declaration closes over constructor parameters. To avoid including all 
of ML, we only pick constructs for working with options and lists that we need later.
%
\begin{equation*}
\begin{array}{rcl}
 \tau &\narrow{=}& \ident{int} \lsep \ident{decimal} \lsep \ident{float} \lsep \ident{bool} \lsep \ident{string} \\[0.0em]
      &\narrow{|}& C \lsep \ident{Data} \lsep \ident{list}\langl\tau\rangl \lsep \ident{option}\langl\tau\rangl \\[0.6em]
 L &\narrow{=}& \kvd{type}~C(\overline{x:\tau}) = \overline{M} \\[0.0em]
 M &\narrow{=}& \kvd{member}~N:\tau=e \\[0.6em]
 v &\narrow{=}& s \lsep \ident{None} \lsep \ident{Some}(v) \lsep \kvd{new}~C(\overline{v}) \lsep [v_1; \ldots; v_n] \\[0.0em]
 e &\narrow{=}& s \lsep op \lsep e.N \lsep \kvd{new}~C(\overline{e}) \lsep {\kvd{if}~e_1~\kvd{then}~e_2~\kvd{else}~e_3}\\
   &\narrow{|}& \ident{None} \lsep\kvd{match}~e~\kvd{with}~\ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2 \\
   &\narrow{|}& \ident{Some}(e) \lsep [e_1;\ldots;e_n]\lsep \ident{map}~(\lambda x\rightarrow e_1)~e_2\\
   &\narrow{|}& e = e \lsep \kvd{match}~e~\kvd{with}~[x_1; \ldots; x_n] \rightarrow e_1 \,|\, \_ \rightarrow e_2
\end{array}
\end{equation*}

\noindent
The type \ident{Data} is a type of all structural values $s$. A class
definition $L$ consists of a constructor and zero or more members. Values $v$ include 
previously defined structural values $s$ and values for the option and list type; finally 
expressions $e$ include previously defined operations $op$, class construction, member access, 
conditionals and expressions for working with option values and lists. We include 
\ident{map} as a special construct to avoid making the language too complex.

\paragraph{Reduction.}
The reduction relation is of the form $e \reduce e'$. We also write 
$e \reduce^{*} e'$ to denote the reflexive and transitive closure of $\reduce$. The reduction rules
for operations $op$ were discussed earlier. Figure~\ref{fig:ff-reduction} shows the remaining
rules.

The (\emph{ctx}) rule performs a reduction inside a sub-expression specified by an evaluation context.
This models the eager evaluation of F\#. An evaluation context $E$ is defined as:
%
\begin{equation*}
\begin{array}{rcl}
 E &\narrow{=}& [\overline{v};E;\overline{e}] \lsep E.N \lsep \kvd{new}~C(\overline{v}, E, \overline{e})\\[0.1em]
   &\narrow{|}& \ident{map}~(\lambda x\rightarrow e)~E \lsep \kvd{if}~E~\kvd{then}~e_1~\kvd{else}~e_2 \\[0.1em]
   &\narrow{|}& \ident{Some}(E) \lsep op(E) \lsep E = e \lsep v = E \\[0.1em]
   &\narrow{|}& \kvd{match}~E~\kvd{with}~\ident{Some}(x) \rightarrow e_1 \,|\, \ident{None} \rightarrow e_2 \\[0.1em]
   &\narrow{|}& \kvd{match}~E~\kvd{with}~[x_1; \ldots; x_n] \rightarrow e_1 \,|\, \_ \rightarrow e_2
\end{array} 
\end{equation*}

\noindent
The evaluation first reduces arguments of functions and the evaluation proceeds from left to right 
as denoted by $\overline{v}, E, \overline{e}$ in constructor arguments or $\overline{v};E;\overline{e}$
in list initialization.

We write $e[\overline{x} \leftarrow \overline{v}]$ for the result of replacing variables $\overline{x}$ by
values $\overline{v}$ in an expression. The (\emph{member}) rule reduces a member access using a class 
definition in the assumption to obtain the body of a member. The remaining six rules
give standard reductions for conditionals and pattern matching.

We omit expressions $e \vee e$ and $e \wedge e$ that are used in Figure~\ref{fig:op-conversions}, 
as those can be easily defined as syntactic sugar using \kvd{if}. All expressions reduce to a value in a 
finite number of steps or get stuck due to an error condition. An error condition can be a wrong 
argument passed to conditional, pattern matching or one of the primitives from Figure~\ref{fig:op-conversions}.

% -------------------------------------------------------------------------------------------------

\begin{figure}
\noindent  
\begin{equation*}
\inference
  {L; \Gamma \vdash e_1 : \ident{Data}}
  {L; \Gamma \vdash [e_1; \ldots; e_n]:\ident{Data}}
\quad  
\inference
  {~}
  {L; \Gamma \vdash n : \ident{int}}
\end{equation*}
\begin{equation*}
\inference
  {L; \Gamma \vdash e_1 : \tau}
  {L; \Gamma \vdash [e_1; \ldots; e_n]:\ident{list}\langl\tau\rangl}
\quad
\inference
  {~}
  {L; \Gamma \vdash s : \ident{Data}}
\end{equation*}
\vspace{0.5em}
\begin{equation*}
\inference
  {L; \Gamma \vdash e : C \\ \kvd{type}~C(\overline{x:\tau}) = ..\;\kvd{member}~N_i : \tau_i = e_i\;.. \in L}
  {L; \Gamma \vdash e.N_i:\tau_i}
\end{equation*}
\vspace{0.25em}
\begin{equation*}
\inference
  {L; \Gamma \vdash e_i : \tau_i & \kvd{type}~C(x_1:\tau_1, \ldots, x_n:\tau_n) = \ldots \in L}
  {L; \Gamma \vdash \kvd{new}~C(e_1, \ldots, e_n):C}
\end{equation*}

\caption{Featherweight F\# -- Fragment of type checking}
\label{fig:ff-typecheck}
\end{figure}

% -------------------------------------------------------------------------------------------------

\paragraph{Type checking.} 
Typing rules in Figure~\ref{fig:ff-typecheck} are written using a judgement
$L; \Gamma \vdash e : \tau$ where the context also contains a set of class declarations $L$.
The fragment demonstrates the key differences from Standard ML and Featherweight Java:
%
\begin{itemize}
\item[--] All structural values $s$ have the type \ident{Data}, but some have other types
  (Booleans, strings, integers, etc.) as illustrated by the rule for $n$.
  For other values, \ident{Data} is the only type -- this includes records and \kvd{null}.
\item[--] A list containing other structural values $[s_1; \ldots; s_n]$ has a type \ident{Data},
  but can also have the $\ident{list}\langl\tau\rangl$ type. Conversely, lists that contain
  non-structural values like objects or options are not of type \ident{Data}.
\item[--] Operations $op$ (omitted) are treated as functions, accepting 
  \ident{Data} and producing an appropriate result type.
\item[--] Rules for checking class construction and member access are similar to corresponding
  rules of Featherweight Java.  
\end{itemize}
%
An important part of Featherweight Java that is omitted here is the checking of type declarations
(ensuring the members are well-typed). We consider only classes generated by our type providers 
and those are well-typed by construction.

% -------------------------------------------------------------------------------------------------

\begin{figure*}
\noindent
\begin{equation*}
\quad\qquad\boxed{
\begin{array}{rcl}
 \nameoftag(\ident{string}) &\narrow{=}& \ident{String} \\
 \nameoftag(\ident{bool}) &\narrow{=}& \ident{Boolean} \\
\end{array}
\qquad
\begin{array}{rcl}
 \nameoftag(\ident{number}) &\narrow{=}& \ident{Number} \\
 \nameoftag(\ident{collection}) &\narrow{=}& \ident{List} \\
\end{array}
\qquad
\begin{array}{rcl}
 \nameoftag(\ident{record}) &\narrow{=}& \ident{Record} \\
 \nameoftag(\ident{named}\;\nu) &\narrow{=}& \nu \\
\end{array}
}
\end{equation*}

\begin{multicols}{2}
% Primitive structural types become corresponding F# types and conversion is inserted
\noindent
\begin{equation*}
\begin{array}{l}
 \sem{\sigma_p}_e = \tau_p,op(e),\emptyset\quad\textnormal{where}~\tau_p = \sigma_p \\[0.2em]
\quad\sigma_p, op\in  \{~(\ident{decimal},\ident{asDec}), (\ident{float},\ident{asFloat})~\}
\end{array}
\end{equation*}
%
% Records become classes
\begin{equation*}
\begin{array}{l}
 \sem{\,\nu_? \{ \nu_1 : \sigma_1, \ldots, \nu_n : \sigma_n \}\,}_e = \\[0.1em]
 \quad C, \kvd{new}~C(e), L_1\cup\ldots\cup L_n\cup\{ L \}\quad\textnormal{where}\\[0.6em]
 \qquad \;\;C~\textnormal{is a fresh class name} \\[0.1em]
 \qquad \;\,\,L = \kvd{type}~C(v:\ident{Data})~=~M_1\ldots M_n  \\[0.1em]
 \qquad M_i = \kvd{member}~\nu_i:\tau_i=e_i\\[0.1em]
 \qquad \tau_i, e_i, L_i = \sem{\sigma_i}_{e'},\;\; e'=\ident{getField}(\nu_?, \nu_i, e)\\[0.6em]
\end{array}
\end{equation*}
%
% Collections
\begin{equation*}
\begin{array}{l}
 \sem{\,[\sigma]\,}_e = \ident{list}\langl\tau\rangl, e_b, L \;\;\textnormal{where}\\[0.4em]
 \qquad e_b = \ident{map}~(\lambda x\rightarrow e')~(\ident{getChildren}(e)) \\[0.1em]
 \qquad \tau, e', L = \sem{\hat{\sigma}}_x
\end{array}
\end{equation*}
%
% Anything else
\begin{equation*}
\begin{array}{l}
 \sem{\bot}_e = \sem{\kvd{null}}_e = \ident{Data}, e, \emptyset
\end{array}
\end{equation*}

\noindent
\begin{equation*}
\hspace{-2.5em}
\begin{array}{l}
 \sem{\sigma_p}_e = \tau_p,e,\emptyset\quad\textnormal{where}~\tau_p = \sigma_p \\[0.2em]
\quad\sigma_p \in  \{~ \ident{string},\ident{bool},\ident{int}~\}\\[0.6em]
\end{array}
\end{equation*}
%
% Sum type
\begin{equation*}
\hspace{-2.5em}
\begin{array}{l}
 \sem{\,\kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl\,}_e = \\[0.1em]
 \quad C, \kvd{new}~C(e), L_1\cup\ldots\cup L_n\cup\{ L \}\;\;\textnormal{where}\\[0.6em]
 \qquad \;\;C~\textnormal{is a fresh class name} \\[0.1em]
 \qquad \;\,\,L = \kvd{type}~C(v:\ident{Data})~=~M_1\ldots M_n \\[0.1em]
 \qquad M_i = \kvd{member}~\nu_i:\ident{option}\langl\tau_i\rangl=\\[0.1em]
 \hspace{5.8em}  \kvd{if}~\ident{hasType}(\sigma_i, v)~\kvd{then}~
     \ident{Some}(e_i)~\kvd{else}~\ident{None} \\[0.1em]
 \qquad \tau_i, e_i, L_i = \sem{\sigma_i}_e,\;\; t_i = \tytagof{(\sigma_i)},\;\; \nu_i=\nameoftag{(t_i)}
\end{array}
\end{equation*}
%
% Option values
\begin{equation*}
\hspace{-2.5em}
\begin{array}{l}
 \sem{\hat{\sigma}\;\,\kvd{option}}_e = \\[0.2em]
 \hspace{1.25em} \ident{option}\langl\tau\rangl, \kvd{if}~\ident{isNull}~e~\kvd{then}~\ident{None}~\kvd{else}~\ident{Some}(e'), L\\[0.2em] 
 \hspace{1.25em} \textnormal{where}~\tau, e', L = \sem{\hat{\sigma}}_e
\end{array}
\end{equation*}
\end{multicols}

\caption{Type provider -- generation of featherweight F\# types from inferred structural types}
\label{fig:tp-generation}
\vspace{-0.5em}
\end{figure*}

% -------------------------------------------------------------------------------------------------

\subsection{Type providers}
\label{sec:formal-tp}

So far, we defined the type inference algorithm which produces a structural type $\sigma$ from one 
or more sample documents (\S\ref{sec:inference}) and we defined a simplified model of evaluation
of F\# (\S\ref{sec:formal-ff}) and F\# Data runtime (\S\ref{sec:formal-tp}). In this section, we 
define how the type providers work, linking the two parts.

All F\# Data type providers take (one or more) sample documents, infer a preferred typing $\sigma$
and then use it to generate F\# types that are exposed to the programmer.\footnote{The actual 
implementation provides \emph{erased types} as described in \cite{fsharp-typeprov}. Here, we treat 
the code as actually generated. This is an acceptable simplification, because F\# Data type providers 
do not rely on laziness that is available through erased types.}

\paragraph{Type provider mapping.}
A type provider produces an F\# type $\tau$ together with an expression that wraps an input
value (of type \ident{Data}) as a value of type $\tau$ and a collection of class definitions. 
We express it using the following mapping:
%
\begin{equation*}
\sem{-}_{-} : (\sigma\times e) \rightarrow (\tau \times e' \times L)
\end{equation*}
%
The mapping $\sem{\sigma}_e$ takes an inferred structural type $\sigma$ and an expression $e$, 
which represents code to obtain a structural value that is being wrapped. It returns an F\# type 
$\tau$, an expression $e'$ which constructs a value of type $\tau$ using $e$ and also a set of 
generated class definitions $L$.

Figure~\ref{fig:tp-generation} shows the rules that define $\sem{-}_{-}$. Primitive types are all
handled by two rules -- for \ident{decimal} and \ident{float}, we insert a call to an appropriate 
conversion function from Figure~\ref{fig:op-conversions}.

For records, we generate a class $C$ that takes a structural value as constructor
parameter. For each record field, we generate a member with the same name as the field.\footnote{The actual
F\# Data implementation also capitalizes the names.} The body of the member calls \ident{getField} and then
wraps this expression using $\sem{\sigma_i}$ which maps the field (structural value of type $\sigma_i$) into 
an F\# value of type $\tau_i$. The returned expression creates a new instance of $C$ and the mapping returns 
the class $C$ together with all recursively generated classes. Note that the class
name $C$ is not directly accessed by the user and so we can use arbitrary name, although the actual 
implementation in F\# Data attempts to infer a reasonable name.\footnote{For example, in 
\ident{\{\str{person}:\{\str{name}:\str{Tomas}\}\}}, the nested record will be named \ident{Person}
based on the name of the parent record field.}

A collection type becomes an F\# $\ident{list}\langl\tau\rangl$. The returned expression calls \ident{getChildren}
(which turns \kvd{null} into an empty list) and then uses \ident{map} to convert nested values to 
values of an F\# type $\tau$. The handling of option type is similar; if the value is not \kvd{null},
we wrap the recursively generated conversion expression $e'$ in the \ident{Some} constructor.

As discussed earlier, variant types are also generated as F\# classes with properties. Given a variant
$\kvd{any}\langl\sigma_1, \ldots, \sigma_n\rangl$, we get corresponding F\# types $\tau_i$ and generate 
$n$ members of type $\ident{option}\langl \tau_i\rangl$. When the member is accessed, we need to perform
a runtime type test using \ident{hasType} to ensure that the value has the right type (similarly to runtime 
type conversions from the top type in languages like Java). If the type matches, a \ident{Some} value is 
returned. The type inference algorithm also guarantees that there is only one case for each type tag 
(\S\ref{sec:inference-commonsuper}) and so we can use the tag when naming the generated member 
(using the \ident{nameof} function).

\paragraph{Example 1.}
To illustrate how the mechanism works, we consider two examples. First, assume 
that the inferred type is a record  
$\{~\ident{Age}\!:\!\kvd{option}\langl\ident{int}\rangl, \ident{Name}\!:\!\ident{string}~ \}$. 
The rules from Figure~\ref{fig:tp-generation} produce the following class:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{Person}(\ident{v}:\ident{Data})~= \\[0.1em]
 \quad \kvd{member}~\ident{Age}~:~\ident{option}\langl\ident{int}\rangl~= \\[0.1em]
 \qquad \kvd{if}~\ident{isNull}~(\ident{getField}(\ident{Person},\ident{Age}, \ident{v}))~\kvd{then}~\ident{None} \\[0.1em]
 \qquad \kvd{else}~\ident{Some}(\ident{asInt}(\ident{getField}(\ident{Person},\ident{Age}, \ident{v}))) \\[0.1em]
 \quad \kvd{member}~\ident{Name}~:~\ident{string}~= \\[0.1em]
 \qquad \ident{asStr}(\ident{getField}(\ident{Person},\ident{Name}, \ident{v}))
\end{array}
\end{equation*}
%
The body of the \ident{Age} member is produced by the case for optional types applied to an expression
$\ident{getField}(\ident{Person},\ident{Age},v)$. If the returned field is not \kvd{null}, then the member
calls \ident{asInt} and wraps the result in \ident{Some}. Note that \ident{getField} is defined even 
when the field does not exist, but it returns \kvd{null}. This lets us treat missing fields as 
optional. An F\# type corresponding to the (structural) record is \ident{Person} and a structural 
value $s$ is wrapped by by calling $\kvd{new}~\ident{Person}(s)$.

\paragraph{Example 2.} The second example illustrates the remaining types including collections and 
variants. Reusing \ident{Person} from the previous example, consider 
$[\kvd{any}\langl\ident{Person},\ident{string}\rangl]$, a collection which contains a mix 
of \ident{Person} and string values:
%
\begin{equation*}
\begin{array}{l}
 \kvd{type}~\ident{PersonOrString}(v:\ident{Data})~= \\[0.1em]
 \quad \kvd{member}~\ident{Person}~:~\ident{option}\langl\ident{Person}\rangl~= \\[0.1em]
 \qquad \kvd{if}~\ident{hasType}(\{\ident{Age}\!:\!\kvd{option}\langl\ident{int}\rangl, \ident{Name}\!:\!\ident{string}\}, v)~\\[0.1em]
 \qquad\quad \kvd{then}~\ident{Some}(\kvd{new}~\ident{Person}(v))~\kvd{else}~\ident{None} \\[0.1em]
 \quad \kvd{member}~\ident{String}~:~\ident{option}\langl\ident{string}\rangl~= \\[0.1em]
 \qquad \kvd{if}~\ident{hasType}(\ident{string}, v)~\\[0.1em]
 \qquad\quad \kvd{then}~\ident{Some}(\ident{asStr}(v))~\kvd{else}~\ident{None}
\end{array}
\end{equation*}
%
The type provider maps the structural type to an F\# type $\ident{list}\langl\ident{PersonOrString}\rangl$. 
Given a structural document value $s$, the code to obtain the wrapped F\# value is:
%
\begin{equation*}
\ident{map}~(\lambda x\rightarrow\kvd{new}~\ident{PersonOrString}(x))~(\ident{getChildren}(s))
\end{equation*}

The \ident{PersonOrString} type contains one member for each of the variant case. In the body, they
check that the value has the correct type using \ident{hasType}. This also implicitly handles \kvd{null}
by returning \kvd{false}. As discussed earlier, variants provide easy access to the known cases
(\ident{string} or \ident{Person}), but they are robust with respect to unexpected inputs.

% --------------------------------------------------------------------------------------------------

% ==================================================================================================
%
%    #######
%       #    #   # #####  ######     ####    ##   ###### ###### ##### #   #
%       #     # #  #    # #         #       #  #  #      #        #    # #
%       #      #   #    # #####      ####  #    # #####  #####    #     #
%       #      #   #####  #              # ###### #      #        #     #
%       #      #   #      #         #    # #    # #      #        #     #
%       #      #   #      ######     ####  #    # #      ######   #     #
%
% ==================================================================================================

\section{Relativized type safety}
\label{sec:safety}

Informally, the safety property for structural type providers states that, given representative sample
documents, any code that can be written using the provided types is guaranteed to work. We call this 
\emph{relativized safety}, because we cannot avoid \emph{all} errors. In particular, one can always
provide an input that has a different structure than any of the samples. In this case, it is expected 
that the code will throw an exception in the implementation (or get stuck in our model).

More formally, given a set of sample documents, code using the provided type is guaranteed to work if 
the inferred type of the input is a subtype of any of the samples. Going back to 
\S\ref{sec:inference-subtyping}, this means that:
%
\begin{itemize}[noitemsep]
\item[--] Input can contain smaller numerical values (e.g., if a sample contains float, the input can contain an integer)
\item[--] Records in the input can have additional fields
\item[--] Records in the input can have fewer fields, provided that the type of the field is missing in some of the samples
\item[--] When a variant is inferred from the sample, the actual input can also contain any other value  
\end{itemize}
%
The following lemma states that the provided code (generated in Figure~\ref{fig:tp-generation})
works correctly on an input $s'$ that is a subtype of the sample $s$. More formally, the provided
provided expression (with input $s'$) can be reduced to a value and, if it is a class,
all its members can also be reduced to values.

\begin{lemma}[Correctness of provided types]
\label{thm:tp-correctness}
Given a sample value $s$ and an input value $s'$ such that $\semalt{s} \succ s'$
and provided type, expression and class declarations $\tau, e, L = \sem{\semalt{s}}_{s'}$, 
then $e \reduce^{*} v$ and if $\tau$ is a class ($\tau=C$) then for all members $N_i$ of the 
class $C$, it holds that $e.N_i \reduce^{*} v$.
\end{lemma}
\begin{proof}
By induction over the structure of $\sem{-}_-$. For primitives, the conversion functions accept all subtypes.
For other cases, analyse the provided code to see that it can work on all subtypes (\emph{e.g.}~\ident{getChildren}
works on \kvd{null} values, \ident{getField} returns \kvd{null} when a field is missing and, for variants,
the value has the correct type as guaranteed by \ident{hasType}.
\end{proof}

\noindent
This shows that the provided types are correct with respect to subtyping. Our key
theorem states that, for any input (which is a subtype of any of the samples) and 
any expression $e$, a well-typed program that uses the provided types does not ``go wrong''.
Using standard syntactic type safety  \cite{syntactic}, we prove type preservation 
(reduction does not change type) and progress (an expression that is not a value can be reduced).

\begin{theorem}[Relativized safety]
\label{thm:safety}
Assume $s_1, \ldots, s_n$ are samples, $\sigma=\semalt{s_1, \ldots, s_n}$ is an inferred
type and $\tau,e,L = \sem{\sigma}_x$ are a type, expression and class definitions generated by a 
type provider.

For all new inputs $s'$ such that $\exists i.(\semalt{s_i} \leq \semalt{s'})$, let $e_s=e[x\leftarrow s']$
be an expression (of type $\tau$) that wraps the input in a provided type. Then, for any expression $e_c$
(user code) such that $\emptyset; y:\tau \vdash e_c:\tau'$ and $e_c$ does not contain primitive operations
$op$ as a sub-expression, it is the case that $e_c[y\leftarrow e_s] \reduce^{*} v$ for some value $v$ and
also $\emptyset; \vdash v : \tau$.
\end{theorem}
\begin{proof}
We discuss the two parts of the proof separately as type preservation (Lemma~\ref{thm:rs-preservation})
and progress (Lemma~\ref{thm:rs-progress}).
\end{proof}

\begin{lemma}[Preservation]
\label{thm:rs-preservation}
Given the class definitions $L$ generated by a type provider as specified in
the assumptions of Theorem~\ref{thm:safety}, then if $\Gamma \vdash e : \tau$ and 
$e \reduce^{*} e'$ then $\Gamma \vdash e' : \tau$.
\end{lemma}
\begin{proof}
By induction over the reduction $\reduce$. The cases for the ML subset of Featherweight F\# 
are standard. For (\emph{member}), we check that code generated by type providers
in Figure~\ref{fig:tp-generation} is well-typed.
\end{proof}

\noindent
The progress lemma states that evaluation of a well-typed program does not reach an undefined state. 
This is not a problem for the (standard) ML subset and object-oriented subset of the calculus. The 
problematic part are the primitive functions (Figure~\ref{fig:op-conversions}). Given a structural 
value (of type \ident{Data}), the reduction can get stuck if the value does not have a structure 
required by a specific primitive.

The Lemma~\ref{thm:tp-correctness} guarantees that this does not happen inside the provided type.
In Theorem~\ref{thm:safety}, we carefully state that we only consider expressions $e_c$ which 
``[do] not contain primitive operations $op$ as sub-expressions''. This makes sure that the only 
code working with structural values is the code generated by the type provider.

\begin{lemma}[Progress]
\label{thm:rs-progress}
Given the assumptions and definitions from Theorem~\ref{thm:safety}, it is the case that
$e_c[y\leftarrow e_s] \reduce e_c'$.
\end{lemma}
\begin{proof}
Proceed by induction over the typing derivation of $L; \emptyset \vdash e_c[y\leftarrow e_s] : \tau'$. 
The cases for the ML subset are standard. For member access, we rely on Lemma~\ref{thm:tp-correctness}.
\end{proof}

% ==================================================================================================
%
%   ###                                                                                     
%    #  #    # #####  #      ###### #    # ###### #    # #####   ##   ##### #  ####  #    # 
%    #  ##  ## #    # #      #      ##  ## #      ##   #   #    #  #    #   # #    # ##   # 
%    #  # ## # #    # #      #####  # ## # #####  # #  #   #   #    #   #   # #    # # #  # 
%    #  #    # #####  #      #      #    # #      #  # #   #   ######   #   # #    # #  # # 
%    #  #    # #      #      #      #    # #      #   ##   #   #    #   #   # #    # #   ## 
%   ### #    # #      ###### ###### #    # ###### #    #   #   #    #   #   #  ####  #    # 
%
% ==================================================================================================

\section{Practical experience}
\label{sec:impl}

The F\# Data library has been widely adopted by the industry and is one of the most downloaded
F\# libraries.\footnote{At the time of writing, the library has over 80,000 downloads on NuGet 
(package repository), 1,821 commits and 44 contributors on GitHub.} A practical demonstration of 
development using the library can be seen in an attached screencast and additional documentation can be
found at \url{http://fsharp.github.io/FSharp.Data}.

In this section, we discuss our practical experience with the safety guarantees provided by the
F\# Data type providers and other notable aspects of the implementation.

\subsection{Relativized safety in practicve}
\label{sec:safety-discuss}

The \emph{relativized safety} property does not guarantee the same amount of safety as traditional
ML type safety, but it reflects the reality of programming with external data sources that is 
increasingly important \cite{age-of-web}. Type providers do not reduce the safety -- they simply 
illuminate the existing issues.

One such issue is the handling of schema change. With type providers, the sample is captured 
at compile-time. If the schema changes later (so that the input is no longer a subtype of the 
sample), the program fails at runtime and developers have to handle the exception. This is the 
same problem that happens when reading data using any other library.

F\# Data can help discover such errors earlier. Our first example (\S\ref{sec:introduction})
points the JSON type provider at a sample using a live URL. This has the advantage that a re-compilation 
fails when the schema changes, which is an indication that the program needs to be updated to reflect the
change.

In general, there is no better solution for plain XML, CSV and JSON data sources. However, some data 
sources provide versioning support (with meta-data about how the schema changed). For those, a type 
provider could adapt automatically, but we leave this for future work.

% -------------------------------------------------------------------------------------------------

\subsection{Parsing structured data}
\label{sec:impl-parsing}

In our formalization, we treat XML, JSON and CSV uniformly as \emph{structural values}. With the addition of 
named records (for XML nodes), the definition of structural values is rich enough to capture all 
three formats\footnote{The same mechanism has later been used by the HTML type provider 
(\url{http://fsharp.github.io/FSharp.Data/HtmlProvider.html}), which provides similarly easy 
access to data in HTML tables and lists.}. However, parsing real-world data poses a number of practical issues.

\paragraph{Reading CSV data.}
When reading CSV data, we read each row as an unnamed record and return a collection of rows.
One difference between JSON and CSV is that in CSV, the literals have no data types and so 
we also need to infer the type of primitive values. For example:
%
{\small{
\begin{verbatim}
  Ozone, Temp, Date,       Autofilled
  41,    67,   2012-05-01, 0
  36.3,  72,   2012-05-02, 1
  12.1,  74,   3 kveten,   0
  17.5,  #N/A, 2012-05-04, 0
\end{verbatim}
}}
%
\noindent
The value {\small\ttfamily \#N/A} is commonly used to represent missing values in CSV and is treated
as \kvd{null}. The \ident{Date} column uses mixed formats and is inferred as \ident{string} 
(we support many date formats and ``May 3'' would be parsed correctly). More interestingly,
we also infer \ident{Autofiled} as Boolean, because the sample contains only $0$ and $1$.
This is handled by adding a \ident{bit} type which is a subtype of both \ident{int} and \ident{bool}.

\paragraph{Reading XML documents.}
Mapping XML documents to structural values is perhaps the most interesting. For each node, we
create a named record. Attributes become record fields and the body becomes a field with a special
name:
%
{\small{
\begin{verbatim}
  <root id="1">
    <item>Hello!</item>
  </root>    
\end{verbatim}
}}
%
\noindent
This XML becomes a record \ident{root} with fields \ident{id} and $\bullet$ for the body. 
The nested element contains only the $\bullet$ field with the inner text. As with CSV, we
infer type of primitive values:
%
\begin{equation*}
\ident{root}~\{ \ident{id} \mapsto 1, \bullet \mapsto [ \ident{item}~\{ \bullet \mapsto \str{Hello!} \}] \}
\end{equation*}
%
When generating types for XML documents, we also lifts the members nested under the $\bullet$ 
field into the parent type to simplify the structure of the generated type.

The XML type provider also includes an option to use \emph{global inference}. In that case, 
the inference from values (\ref{sec:formal-inferval}) unifies the types of all records with the 
same name. This is useful because, for example, in XHTML all {\small\ttfamily <table>} elements
will be treated as values of the same type.

% -------------------------------------------------------------------------------------------------

\subsection{Heterogeneous collections} 
\label{sec:impl-hetero}

When introducing type providers (\S\ref{sec:providers-sum}), we mentioned that F\# Data 
implements a special handling of heterogeneous collections. This allows us to avoid inferring a 
variant types in many common scenarios. In the earlier example, a sample collection 
contains a record (with \strf{pages} and \strf{page} fields) and a nested collection with values.

Rather than storing a single type for the collection elements as in $[\sigma]$, heterogeneous
collections store multiple possible element types together with their \emph{  inferred multiplicity} 
(zero or one, exactly one, zero or more):
%
\begin{equation*}
\begin{array}{rcl}
 \psi &\narrow{=}& 1? \lsep 1 \lsep \ast \\
 \sigma &\narrow{=}& ~\ldots \lsep [\sigma_1, \psi_1 | \ldots | \sigma_n, \psi_n ]
\end{array}
\end{equation*}
%
We omit the details, but finding a preferred typing of two heterogeneous 
collections is analogous to the handling of variants. We merge cases with the same tag (by finding 
their common super-type) and calculate their new shared multiplicity (for example, by turning 
$1$ and $1?$ into $1?$).

% ==================================================================================================
%
%    ######
%    #     # ###### #        ##   ##### ###### #####     #    #  ####  #####  #    #
%    #     # #      #       #  #    #   #      #    #    #    # #    # #    # #   #
%    ######  #####  #      #    #   #   #####  #    #    #    # #    # #    # ####
%    #   #   #      #      ######   #   #      #    #    # ## # #    # #####  #  #
%    #    #  #      #      #    #   #   #      #    #    ##  ## #    # #   #  #   #
%    #     # ###### ###### #    #   #   ###### #####     #    #  ####  #    # #    #
%
% ==================================================================================================

\section{Related and future work}
\label{sec:related}

We connect two lines of research: integration of external data into statically-typed 
programming languages and inferring types for real-world data sources. We build on F\# type 
providers \cite{fsharp-typeprov,fsharp-typeprov-ddfp,idris-tp,liteq}, but our paper is novel 
in that it discusses the theory and safety of a concrete type provider.

\paragraph{Extending the type systems.} 
Previous work that integrates external data, such as XML \cite{xduce,xduce-ml} and databases 
\cite{links,linq}, into a programming language, required the user to define the schema
or it has an ad-hoc extension that reads the schema.

C$\omega$ \cite{comega-xs} is the most similar to F\# Data. It extends C\# with types similar to our
structural types (including similar heterogeneous collections), but it does not infer the types
from a sample and extends the type system of the host language (rather than using general purpose
embedding).

\paragraph{Advanced type systems.}
A number of other advanced type system features could be used to tackle the problem discussed 
in this paper. The Ur \cite{ur} language has a rich system for working with records; 
meta-programming \cite{template-hask}, \cite{th-camlp4} and multi-stage programming \cite{multi-stage}
could be used to generate code for the provided types; and gradual typing \cite{gradual,gradual-js} 
can add typing to existing dynamic languages. As far as we are aware, none of these 
systems have been used to provide the same level of integration with XML, CSV and JSON.

\paragraph{Typing real-world data.}
Recent work \cite{typing-json} infers a succinct type of large JSON datasets using MapReduce.
It fuses similar types based on similarity. This is more sophisticated than our technique, but it 
makes formal specification of safety (Theorem~\ref{thm:safety}) difficult. Extending our 
\emph{relativized safety} to \emph{probabilistic safety} is an interesting future work.

The PADS project \cite{pads-dsl,pads-ml} tackles a more general problem of handling \emph{any} data format.
The schema definitions in PADS are similar to our structural type. The structure inference for LearnPADS
\cite{pads-learn} infers the data format from a flat input stream. A PADS type provider could follow
many of the patterns we explore in this paper, but formally specifying the safety property would be
challenging.

\section{Conclusions}
\label{sec:conclusions}

We explored the F\# Data type providers for XML, CSV and JSON. As most real-world data do not come 
with an explicit schema, the library uses \emph{type inference} that deduces a type from a set of 
samples. Our type inference algorithm is based on a preferred typing relation. For usability 
reasons, it prefers records and avoids variant types. The algorithm is predictable, which is 
important as developers need to understand how changing the samples affects the resulting types.

We explore the programming language theory behind type providers. F\# Data is a prime example of 
type providers, but our work also demonstrates a more general point. The types generated by type 
providers can depend on external input and so we can only guarantee \emph{relativized safety}, 
which says that a program is safe only if the actual inputs satisfy additional conditions --
in our case, they have to be subtypes of one of the samples.

Type providers have been described before, but this paper is novel in that it explores concrete 
type providers from the perspective of programming language theory. This is particularly important
for F\# Data type providers, which have been widely adopted by the industry.

% \acks
% We would like to thank to the F\# Data contributors on GitHub and other colleagues working on type providers,
% including Jomo Fisher, Keith Battocchi and Kenji Takeda.

\bibliographystyle{abbrvnat}
\bibliography{paper}

\end{document}